Using sparseHessianFD

### 2015-01-09 inst/doc/sparseHessianFD.Rmd
version 0.2.0 probably ready for CRAN

```{r, echo=FALSE}
N <- 6
k <- 2
nv1 <- (N+1)*k
nels1 <- nv1^2
nnz1 <- (N+1)*k^2 + 2*N*k^2
nnz1LT <- (N+1)*k*(k+1)/2 + N*k^2
Q <- 1000
nv2 <- (Q+1)*k
nels2 <- nv2^2
nnz2 <- (Q+1)*k^2 + 2*Q*k^2
nnz2LT <- (Q+1)*k*(k+1)/2 + Q*k^2
options(scipen=999)
```

@@

The functions for computing the objective function, gradient and
Hessian for this example are in the R/ex\_funcs.R file.  The package
also includes a sample dataset with simulated data from the binary choice example.

@@

```{r}
set.seed(123)
data(binary)
Y <- binary$Y
X <- binary$X
T <- binary$T
N <- length(Y)
k <- NROW(X)
nvars <- as.integer(N*k + k)
P <- rnorm(nvars) ## random starting values
inv.Sigma <- rWishart(1,k+5,diag(k))[,,1]
inv.Omega <- diag(k)
```

@@

The objective function for the binary choice example is `binary.f` and the gradient function is
`binary.grad`.  The first argument to both is the variable vector, and
the argument lists must be the same for both.  For this example, we
need to provide the data ($X$ and $Y$) and the prior parameters
($\Sigma^{-1}$ and $\Omega^{-1}$). The functions also have an
`order.row` argument to change the ordering of the variables (and
thus, the sparsity pattern). If `order.row=TRUE`, then the Hessian
will have an off-diagonal pattern.  If `order.row=FALSE`, then the
Hessian will have a block-arrow pattern.

@@

```{r, tidy=TRUE}
true.f <- binary.f(P, Y=Y, X=X, T=T, inv.Sigma=inv.Sigma, inv.Omega=inv.Omega, order.row=FALSE)
true.grad <- binary.grad(P, Y=Y, X=X, T=T, inv.Sigma=inv.Sigma, inv.Omega=inv.Omega, order.row=FALSE)
true.hess <- binary.hess(P, Y=Y, X=X, T=T, inv.Sigma=inv.Sigma, inv.Omega=inv.Omega, order.row=FALSE)
```

@@

Next, we create a new instance of a sparseHessianFD object with an
"initial variable" $P$, and the row and column indices of the non-zero
elements in the lower triangle of the Hessian.  We also pass in any
other arguments for `binary.f` and `binary.grad`.  We accept the default
values for other arguments to `sparseHessianFD.new`.
```{r}
obj <- sparseHessianFD.new(P, binary.f, binary.grad,
       rows=pattern$iRow, cols=pattern$jCol,
       X=X, Y=Y, T=T, inv.Sigma=inv.Sigma, inv.Omega=inv.Omega,
       order.row=FALSE)
```

@@

>>>>

```{r, echo=FALSE}
require(Matrix)
require(sparseHessianFD)
N <- 6
k <- 2
nv1 <- (N+1)*k
nels1 <- nv1^2
nnz1 <- (N+1)*k^2 + 2*N*k^2
nnz1LT <- (N+1)*k*(k+1)/2 + N*k^2
Q <- 1000
nv2 <- (Q+1)*k
nels2 <- nv2^2
nnz2 <- (Q+1)*k^2 + 2*Q*k^2
nnz2LT <- (Q+1)*k*(k+1)/2 + Q*k^2
options(scipen=999)
```

@@

The functions for computing the objective function, gradient and
Hessian for this example are in the R/binary.R file.  The package
also includes a sample dataset with simulated data from the binary choice example.

@@

```{r}
set.seed(123)
data(binary)
str(binary)
N <- length(binary$Y)
k <- NROW(binary$X)
nvars <- as.integer(N*k + k)
P <- rnorm(nvars) ## random starting values
priors <- list(inv.Sigma = rWishart(1,k+5,diag(k))[,,1],
               inv.Omega = diag(k))

```

@@

The objective function for the binary choice example is `binary.f` and the gradient function is
`binary.grad`.  The first argument to both is the variable vector, and
the argument lists must be the same for both.  For this example, we
need to provide the data list ($X$, $Y$ and $T$) and the prior
parameter list
($\Sigma^{-1}$ and $\Omega^{-1}$). The functions also have an
`order.row` argument to change the ordering of the variables (and
thus, the sparsity pattern). If `order.row=TRUE`, then the Hessian
will have an off-diagonal pattern.  If `order.row=FALSE`, then the
Hessian will have a block-arrow pattern.

@@

```{r, tidy=TRUE}
true.f <- binary.f(P, binary, priors, order.row=FALSE)
true.grad <- binary.grad(P, binary, priors, order.row=FALSE)
true.hess <- binary.hess(P, binary, priors, order.row=FALSE)
```

@@

Next, we create a new instance of a sparseHessianFD object with an
"initial variable" $P$, and the row and column indices of the non-zero
elements in the lower triangle of the Hessian.  We also pass in any
other arguments for `binary.f` and `binary.grad`.  We accept the default
values for other arguments to `sparseHessianFD.new`.
```{r}
obj <- sparseHessianFD.new(P, binary.f, binary.grad,
       rows=pattern$rows, cols=pattern$cols,
       data=binary, priors=priors,
       order.row=FALSE)
```

@@

## Quick summary
In short, to use the package, follow the following steps:

1.  Write R functions to return the value and gradient of the
objective function.
2.  Determine the row and column indices of the non-zero elements of
the lower triangle of the Hessian.
3.  Pick a variable vector (i.e., a starting value) at which you can
    initialize the sparseHessian object.  It doesn't really matter
    what this vector is, as long as the function value and gradient
    elements are all finite.
3.  Create a new sparseHessianFD object using the sparseHessianFD.new
function.  For this example, call that object F.
4.  Compute Hessian at x by calling F$hessian(x).
5.  ???
6.  Profit.
